Learning Action Models from State Observations

### 2018-01-25 doc/ijcai-18/observations-learning.tex
updated section 5 ijcai18 paper

\section{Evaluation}
This section presents a two-fold evaluation of the learned models: with respect to a reference model and with respect to given sets of observations.

\subsubsection{Reproducibility}
We used 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}. For the learning of the \strips\ action models we only used observations sequences $\tup{s_0, s_1, \ldots, s_{24}}$ of twenty five states per domain. These sets of learning examples are fixed for all the experiments so the results reported by the different evaluation approaches is comparable. All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.

@@

The compilation source code, the evaluation scripts and the benchmarks (including learning and testing sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.

@@

\subsection{Evaluating with a reference model}
Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC. For each domain, the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics:
\begin{itemize}
\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear). Gives a notion of the {\em soundness} of the learned models.
\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing). Gives a notion of the {\em completeness} of the learned models.
\end{itemize}

When the learning hypothesis space is under constrained, the learned actions can be reformulated and still be compliant with the inputs. Given the syntax-based nature of the {\em precision} and {\em recall} metrics, they may report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e. a learned model semantically equivalent but syntactically different to the reference model. To address this issue and give insight to the quality of the learned models we define a more robust method for computing {\em Precision} and {\em Recall} with regard to a reference model.

Precision and recall are often combined using the {\em harmonic mean}. This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$. Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ such that $f_{P\&R}$ maximizes the accumulated {\em F-measure}. Thus each learned action scheme $\xi\in\Xi$ is compared with the scheme $f_{P\&R}(\xi)\in \Xi^*$ indicated by the $f_{P\&R}$ mapping. Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing with the reference scheme indicated by the $f_{P\&R}$ mapping. This shows that in many domains, (almost all of them except the {\em ferry} and {\em satellite} domains) the learned actions models, or their parameters, were learned properly but interchanged their roles with respect to the reference model.

\subsection{Evaluating with a test set}
When a reference action model is not available, then the evaluation of the quality of the learning process can be done with respect to a test set. The difficulty here comes from the fact that how the observations of the testing set are generated strongly affect to the reported results. This is not trivial when learning planing action models since the generation of relevant and random observations of planning domains is an open issue. Planning domains are highly structured which makes that some states are likely to have low probability of being chosen by chance~\cite{fern2004learning}.

Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observations test set. The table reports for each domain, the {\em observation edit distance}computed with our extended compilation, the {\em maximum edit distance} and their ratio.

Interestingly the Table~\ref{fig:observationstest} shows that using a test set results in reporting higher learning scores. The explanation for this is that how the observations are generated affect to the reported results. For instance in the floortile domain the learned action missed learning the action scheme for the {\tt paint-down} action because this action was not considered in the state observations of the learning test. If this same action is not used either to generate the observations of the test set it means that the learned action model will not be penalized for not learning this action scheme.

>>>>

\section{Evaluation}
This section reports a two-fold evaluation of the learned models: (1) with respect to a given observation test set and (2), with respect to a reference model.

\subsubsection{Reproducibility}
We used 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}. For the learning of the \strips\ action models we only used observations sequences $\tup{s_0, s_1, \ldots, s_{24}}$ of twenty five states per domain. The set of learning examples is fixed for all the experiments so the results reported by the different evaluation approaches is comparable. All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.

@@

The compilation source code, the evaluation scripts and the benchmarks (including the learning and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.

\subsubsection{Evaluating with a test set}
When a reference model is not available, the evaluation of the learning process can be done with respect to a test set. Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observation test set. The table reports, for each domain, the {\em observation edit distance} (which is computed with our extended compilation), the {\em maximum edit distance}, and their ratio.

The reported results showed that, despite we are learning only from 25 state observations, twelve out of fifteen learned domains achieved ratios of $90\%$ or above. This fact evidences that the learned models required small edition amounts to match the observations given in the test set.

\subsubsection{Evaluating with a reference model}
Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC. For each domain, the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics:
\begin{itemize}
\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates in the learned model that should not appear). Gives a notion of {\em soundness}.
\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). Gives a notion of {\em completeness}.
\end{itemize}

The learning scores reported in Table~\ref{fig:observationsnomap} are poorer than in Table~\ref{fig:observationstest}. {\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes of the roles of actions or their parameters).

@@

\subsubsection{Precision and recall robust to model reformulations}
To give insight to the actual quality of the learned models we define a method for computing {\em Precision} and {\em Recall} that is robust to such model reformulations. Precision and recall are often combined using the {\em harmonic mean}. This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$. Given the learned \strips\ action model $\Xi$ and the reference \strips\ action model $\Xi^*$, the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.

Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\Xi$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \Xi^*$ given by the $f_{P\&R}$ mapping. These results show that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions (or their parameters) with respect to their role in the reference model.

However in few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}. The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}. For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set, the same happened in the {\em floortile} with the {\tt paint-down} action or in the {\em parking} domain with the {\tt move-curb-to-curb} action. We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized for not learning these action schemes.  The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
